```{r setup, include=FALSE}

rm(list=ls())

library(openxlsx)
library(dplyr)
library(data.table)
library(forecast)
library(ggplot2)
library(knitr)
library(kableExtra)
library(grid)
library(gridExtra)
library(scales)
library(RColorBrewer)
library(ggthemes)

theme_set(theme_minimal())
ret <- function(x,r=0.3,l=0){
  out <- x
  for(i in 2:length(x)){
    out[i] <- out[i-1]*r+x[i]
  }
  out <- c(rep(0,l),out)[1:length(out)]
  out * sum(x)/sum(out)
}
rsq <- function(x,y){
  summary(lm(y~x))$r.square
}
scalelim <- function(x,f=1.5){
  x <- as.numeric(unlist(x))
  (range(x,na.rm=T)-mean(x,na.rm=T))*f+mean(x,na.rm=T)
}
comval <- function(x){
  x <- round(x,1)
  x <- strsplit(paste(x),'')[[1]]
  x[(length(x):1)%%3==0] <- paste0(',',x[(length(x):1)%%3==0])
  x <- paste(x,collapse='')
  if(substr(x,1,1)==','){x <- substr(x,2,nchar(x))}
  return(x)
}
vplayout <- function(x,y){viewport(layout.pos.row = x, layout.pos.col = y)}
temp <- openxlsx::read.xlsx('/Users/wenrurumon/Documents/GSK/forecasting/Fenbid_workingfile.xlsx')
b <- read.csv('/Users/wenrurumon/Documents/GSK/forecasting/coef_summary.csv',row.names=1)
temp$Month <- as.Date(temp$Month)
X <- data.table(
  select(temp,1,2),
  intercept = 1,
  select(temp,5,6,7),
  dummy.vc = ifelse(temp$Month>='2019-4-1',temp$Weighted.Distribution,0),
  select(temp,8:13)
)
for(i in 8:13){X[[i]] <- ret(X[[i]])}
decomp <- as.matrix(select(X,-1:-2))
for(i in 1:ncol(decomp)){decomp[,i] <- decomp[,i] * unlist(b)[i]}
decomp <- as.data.table(decomp)
decomp$Weighted.Distribution <- decomp$Weighted.Distribution + decomp$dummy.vc
decomp <- select(decomp,-dummy.vc)
vplayout <- function(x,y){viewport(layout.pos.row = x, layout.pos.col = y)}

```

---

# Fenbid Forecasting Report

**Nielsen Advanced Analytic Consulting**

`r Sys.Date()`

---

## Content


**1 Result Summary**

**2 Prediction Review**

-- 2.1 Sales Volume

-- 2.2 Sales Volume Growth

-- 2.3 Sales Value

-- 2.4 Sales Value Growth

**3 Source of Sales Change**

-- 3.1 Source of Sales Change

-- 3.2 Monthly Source of Sales Change

**4 Tech Review**

<br/>

---

### 1. Result Summary

<br/>

```{r echo=FALSE, warning=FALSE, message=FALSE}
options(digits=4,scipen=200)
X.summary <- t(data.frame(Month=temp$Month,
           Volume=rowSums(decomp),
           Value=rowSums(decomp)*temp$DU.Price) %>% group_by(year=as.numeric(year(Month))) %>%
  summarise(Volume=sum(Volume),Value=sum(Value)) %>% filter(year>2018))[-1,] %>% as.data.frame
X.summary <- data.frame(
  sapply(X.summary[,1],comval),
  sapply(X.summary[,2],comval),
  paste0(round((X.summary[,2]/X.summary[,1]-1)*100,1),'%'))
colnames(X.summary) <- c(2019,2020,'CH%')
rownames(X.summary) <- c("Volume",'Value')

kable(X.summary)%>%kable_styling(bootstrap_options = c("striped", "hover"))
```

<br/>

---

<br/>

### 2. Prediction Review

<br/>

#### 2.1 Sales Volume


```{r echo=FALSE, warning=FALSE, message=FALSE}
p <- data.frame(month=temp$Month,Actual=temp$Fenbid.DU,Predict=rowSums(decomp)) %>% mutate(
  Predict=ifelse(is.na(Actual),Predict,NA)
) %>% melt(id='month')
p1 <- ggplot() +
  geom_line(data=p,aes(x=month,y=value,colour=variable),size=1) +
  ylim(scalelim(p$value))+
  labs(x='Month',y='Volume (MDU)',colour='')+ theme(legend.position="top")
p0 <- p %>% group_by(year=year(month),month=month(month)) %>% summarise(value=sum(value,na.rm=T)) 
p <- p %>% group_by(month=year(month),variable) %>% summarise(value=sum(value,na.rm=T)/1000)
p2 <- ggplot() +
  geom_bar(data=p,aes(x=month,y=value,fill=variable),stat='identity',
           position=position_stack(reverse=T)) +
  labs(x='Year',y='Volume (MMDU)',fill='')+ theme(legend.position="none")
p <- p %>% group_by(month) %>% summarise(value=sum(value,na.rm=T))
```

`r paste0('Sales Volume (MMDU) for Fenbid is expected at about ',round(p$value[4],2),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`

```{r echo=FALSE, warning=FALSE, message=FALSE}
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))
print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

kable(reshape2::acast(filter(p0,year>2018),year~month))%>%kable_styling(bootstrap_options = c("striped", "hover"))

```

<br/>

<br/>

#### 2.2 Sales Volume Growth


```{r echo=FALSE, warning=FALSE, message=FALSE}
p <- data.frame(month=temp$Month,Actual=temp$Fenbid.DU,Predict=rowSums(decomp)) %>% mutate(
  Predict=ifelse(is.na(Actual),Predict,NA)
) %>% melt(id='month') %>% filter(!is.na(value))
p <- data.frame(p[-1:-12,-3],value=p$value[-1:-12]/p$value[-(nrow(p)+0:-11)]*100-100)
p0 <- p
p1 <- ggplot() +
  geom_bar(data=p,aes(x=month,y=value,fill=variable),size=1,stat='identity') +
  ylim(scalelim(p$value))+
  labs(x='Month',y='Volume Growth (%)',fill='')+ theme(legend.position="top")

p <- data.frame(month=temp$Month,Actual=temp$Fenbid.DU,Predict=rowSums(decomp)) %>% mutate(
  Predict=ifelse(is.na(Actual),Predict,NA)
) %>% melt(id='month') %>% filter(!is.na(value)) %>% group_by(year=year(month)) %>% summarise(value=sum(value,na.rm=T))
p <- data.frame(year=p$year[-1],value=p$value[-1]/p$value[-(nrow(p))]*100-100)
p2 <- ggplot() +
  geom_bar(data=p,aes(x=year,y=value,fill=c(rep('1',2),'2')),stat='identity') +
  labs(x='Year',y='Volume Growth (%)')+ theme(legend.position="none") +
  ylim(scalelim(p$value))
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))
print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

kable(reshape2::acast(filter(p0%>%mutate(year=year(month),month=month(month)),year>2018),year~month))%>%kable_styling(bootstrap_options = c("striped", "hover"))

```

<br/>

<br/>

#### 2.3 Sales Value


```{r echo=FALSE, warning=FALSE, message=FALSE}
p <- data.frame(month=temp$Month,Actual=temp$Fenbid.Value,Predict=rowSums(decomp)*temp$DU.Price) %>% mutate(
  Predict=ifelse(is.na(Actual),Predict,NA)
) %>% melt(id='month')
p1 <- ggplot() +
  geom_line(data=p,aes(x=month,y=value,colour=variable),size=1) +
  ylim(scalelim(p$value))+
  labs(x='Month',y='Value (MRMB)',colour='')+ theme(legend.position="top")
p0 <- p %>% group_by(year=year(month),month=month(month)) %>% summarise(value=sum(value,na.rm=T)) 
p <- p %>% group_by(month=year(month),variable) %>% summarise(value=sum(value,na.rm=T)/1000)
p2 <- ggplot() +
  geom_bar(data=p,aes(x=month,y=value,fill=variable),stat='identity',
           position=position_stack(reverse=T)) +
  labs(x='Year',y='Value (MMRMB)',fill='') + theme(legend.position="none")
p <- p %>% group_by(month) %>% summarise(value=sum(value,na.rm=T))
```

`r paste0('Sales Value (MMRMB) for Fenbid is expected at about ',round(p$value[4],2),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`

```{r echo=FALSE, warning=FALSE, message=FALSE}
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))
print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

kable(reshape2::acast(filter(p0,year>2018),year~month))%>%kable_styling(bootstrap_options = c("striped", "hover"))

```

<br/>

<br/>

#### 2.4 Sales Value Growth


```{r echo=FALSE, warning=FALSE, message=FALSE}
p <- data.frame(month=temp$Month,Actual=temp$Fenbid.Value,Predict=rowSums(decomp)*temp$DU.Price) %>% mutate(
  Predict=ifelse(is.na(Actual),Predict,NA)
) %>% melt(id='month') %>% filter(!is.na(value))
p <- data.frame(p[-1:-12,-3],value=p$value[-1:-12]/p$value[-(nrow(p)+0:-11)]*100-100)
p0 <- p
p1 <- ggplot() +
  geom_bar(data=p,aes(x=month,y=value,fill=variable),size=1,stat='identity') +
  ylim(scalelim(p$value))+
  labs(x='Month',y='Value Growth (%)',fill='')+ theme(legend.position="top")

p <- data.frame(month=temp$Month,Actual=temp$Fenbid.Value,Predict=rowSums(decomp)*temp$DU.Price) %>% mutate(
  Predict=ifelse(is.na(Actual),Predict,NA)
) %>% melt(id='month') %>% filter(!is.na(value)) %>% group_by(year=year(month)) %>% summarise(value=sum(value,na.rm=T))
p <- data.frame(year=p$year[-1],value=p$value[-1]/p$value[-(nrow(p))]*100-100)
p2 <- ggplot() +
  geom_bar(data=p,aes(x=year,y=value,fill=c(rep('1',2),'2')),stat='identity') +
  labs(x='Year',y='Value Growth (%)')+ theme(legend.position="none") +
  ylim(scalelim(p$value))
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))
print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

kable(reshape2::acast(filter(p0%>%mutate(year=year(month),month=month(month)),year>2018),year~month))%>%kable_styling(bootstrap_options = c("striped", "hover"))

```

<br/>

---

<br/>

### 3. Source of Sales Change

<br/>

#### 3.1 Source of Volume Change


```{r echo=FALSE, warning=FALSE, message=FALSE}

p <- data.frame(year=year(temp$Month),decomp) %>% filter(year>2018) 
p <- data.frame(due=(apply(p[,-1],2,function(x){diff(tapply(x,p$year,sum))})/sum(temp$Fenbid.DU[25:36]))[-1])
p <- data.frame(bucket=rownames(p),due=p) %>% melt()
p$variable <- c(rep('Base',3),rep('ATL',4),rep('BTL',2))
p$bucket <- gsub('\\.',' ',p$bucket)
p$bucket <- gsub(' Spending','',p$bucket)
p1 <- p
p <- data.frame(year=year(temp$Month),volume=rowSums(decomp),
                value=rowSums(decomp)*temp$DU.Price) %>% filter(year>2018) %>% group_by(year) %>%
  summarise(Volume=sum(volume),Value=sum(value)) %>% select(2:3) %>% apply(2,diff)
p <- p/c(sum(temp$Fenbid.DU[25:36]),sum(temp$Fenbid.Value[25:36]))
p <- rbind(Value=p[2],Price=diff(p),Volume=p[1],' '=0) %>% melt
colnames(p) <- colnames(p1)
p <- rbind(p,p1)
p2 <- p[nrow(p):1,]
p2$bucket <- factor(p2$bucket,p2$bucket)
p1 <- p1 %>% group_by(variable) %>% summarise(value=sum(value)*100)

ggplot() +
  geom_bar(data=p2,aes(x=bucket,y=value*100,fill=(value>=0)),stat='identity') +
  coord_flip() +
  ylim(max(abs(p$value*100))*c(-1.2,1.2)) +
  labs(x='',y='Source of Volume Change(%)') + 
  theme(legend.position="none")
```

<br/>
**Supporting Data**

```{r echo=FALSE, warning=FALSE, message=FALSE}

temp2 <- temp
for(i in 3:ncol(temp2)){
  if(mean(temp2[[i]],na.rm=T)>10000000){
    temp2[[i]] <- temp2[[i]]/1000000
    colnames(temp2)[i] <- paste0(colnames(temp2)[i],'(MM)')
  } else if(mean(temp2[[i]],na.rm=T)>10000){
        temp2[[i]] <- temp2[[i]]/1000
    colnames(temp2)[i] <- paste0(colnames(temp2)[i],'(M)')
  }
}
X2 <- data.table(
  select(temp2,1,2),
  select(temp2,5,6,7),
  select(temp2,8:13)
)
X.summary <- melt(select(X2,-Status),id='Month') %>%
  group_by(year=year(Month),variable) %>% 
  filter(year>2018) %>%
  summarise(value=sum(value)) %>% arrange(variable) %>%
  mutate(value=ifelse(variable%in%c('Weighted.Distribution','DU.Price'),value/12,value)) %>%
  reshape2::acast(variable~year) %>% as.data.frame
X.summary$CH <- paste0(round((X.summary[,2]/X.summary[,1]-1)*100),'%')
X.summary[,1] <- sapply(X.summary[,1],comval)
X.summary[,2] <- sapply(X.summary[,2],comval)
kable(X.summary)%>%kable_styling(bootstrap_options = c("striped", "hover"))

```

<br/>

<br/>

#### 3.2 Monthly Source of Volume Change


```{r echo=FALSE, warning=FALSE, message=FALSE}

p <- data.frame(month=temp$Month[-1:-12],
  (decomp[-1:-12,]-decomp[-(nrow(decomp)-0:11),])/rowSums(decomp)[-(nrow(decomp)-0:11)]) %>%
  melt(id='month') %>% filter(variable!='intercept') %>%
  mutate(value=value*100,variable=gsub('\\.',' ',variable))
p$variable <- gsub(' Spending','',p$variable)
p$variable <- factor(p$variable,unique(p$variable))
p <- p %>% group_by(month,variable) %>% summarise(value=sum(value))
ggplot() +
  geom_bar(data=p,aes(x=month,y=value,fill=variable),stat='identity') +
  ylim(range(p$value)*2) +
  labs(x='Month',y='Source of Volume Change %',fill='') +
  theme(legend.position="top")
```

<br/>

```{r echo=FALSE, warning=FALSE, message=FALSE}
p <- p %>% mutate(Year=year(month),Month=month(month)) %>% filter(Year>2019) %>% mutate(
  value=paste0(round(value,1),'%')
) %>% reshape2::acast(
  variable~Month
)
kable(p)%>%kable_styling(bootstrap_options = c("striped", "hover"))

```

<br/>

---

### 4. Tech Review

<br/>

```{r echo=FALSE, warning=FALSE, message=FALSE}
p <- data.frame(month=temp$Month,Actual=temp$Fenbid.DU,Fit=rowSums(decomp)) %>% filter(!is.na(Actual))
ggplot() +
  geom_line(data=melt(p,id='month'),aes(x=month,y=value,colour=variable),size=1) +
  ylim(scalelim(p[,2])) +
  labs(
    title='Fit Chart in Sales Volume',
    subtitle=
         paste0('R square: ',round(rsq(p[,3],p[,2]) * 100,2),'%\nMAPE: ',round(mean(abs(p[,3]/p[,2]-1))*100,2),'%'),
    x='Month',y='Volume (MDU)',colour='')
```
<br/>
```{r echo=FALSE, warning=FALSE, message=FALSE}

p <- data.frame(month=p$month[-1:-12],(p[-1:-12,-1]/p[-(nrow(p)-0:11),-1]-1)*100)
ggplot() +
  geom_line(data=melt(p,id='month'),aes(x=month,y=value,colour=variable),size=1) +
  ylim(scalelim(p[,-1])) +
  labs(
    title='Fit Chart in Volume Growth%',
    subtitle=
      paste0('R square: ',round(rsq(p[,3],p[,2]) * 100,2),'%\nMAPE: ',round(mean(abs(p[,3]-p[,2]-1)),2),'%'),
    x='Month',y='Volume Growth (%)',colour='')
```
<br/>
```{r echo=FALSE, warning=FALSE, message=FALSE}

p <- data.frame(month=temp$Month,Actual=temp$Fenbid.Value,Fit=rowSums(decomp)*temp$DU.Price) %>% filter(!is.na(Actual))
ggplot() +
  geom_line(data=melt(p,id='month'),aes(x=month,y=value,colour=variable),size=1) +
  ylim(scalelim(p[,-1])) +
  labs(
    title='Fit Chart in Sales Value',
    subtitle=
      paste0('R square: ',round(rsq(p[,3],p[,2]) * 100,2),'%\nMAPE: ',round(mean(abs(p[,3]/p[,2]-1))*100,2),'%'),
    x='Month',y='Volume (MRMB)',colour='')
```
<br/>
```{r echo=FALSE, warning=FALSE, message=FALSE}

p <- data.frame(month=p$month[-1:-12],(p[-1:-12,-1]/p[-(nrow(p)-0:11),-1]-1)*100)
ggplot() +
  geom_line(data=melt(p,id='month'),aes(x=month,y=value,colour=variable),size=1) +
  ylim(scalelim(p[,-1])) +
  labs(
    title='Fit Chart in Value Growth%',
    subtitle=
      paste0('R square: ',round(rsq(p[,3],p[,2]) * 100,2),'%\nMAPE: ',round(mean(abs(p[,3]-p[,2]-1)),2),'%'),
    x='Month',y='Value Growth (%)',colour='')
```

---

#### END of Prediction Review

