```{r setup, include=FALSE}

rm(list=ls())

library(openxlsx)
library(dplyr)
library(data.table)
library(forecast)
library(ggplot2)
library(knitr)
library(kableExtra)
library(grid)
library(gridExtra)
library(scales)
library(RColorBrewer)
library(ggthemes)

theme_set(theme_minimal())
ret <- function(x,r=0.3,l=0){
  out <- x
  for(i in 2:length(x)){
    out[i] <- out[i-1]*r+x[i]
  }
  out <- c(rep(0,l),out)[1:length(out)]
  out * sum(x)/sum(out)
}
rsq <- function(x,y){
  summary(lm(y~x))$r.square
}
scalelim <- function(x,f=1.5){
  x <- as.numeric(unlist(x))
  (range(x,na.rm=T)-mean(x,na.rm=T))*f+mean(x,na.rm=T)
}
comval <- function(x){
  x <- round(x,0)
  x <- strsplit(paste(x),'')[[1]]
  x[(length(x):1)%%3==0] <- paste0(',',x[(length(x):1)%%3==0])
  x <- paste(x,collapse='')
  if(substr(x,1,1)==','){x <- substr(x,2,nchar(x))}
  return(x)
}
vplayout <- function(x,y){viewport(layout.pos.row = x, layout.pos.col = y)}
temp <- openxlsx::read.xlsx('/Users/wenrurumon/Documents/GSK/forecasting/Fenbid_workingfile.xlsx')
temp$Month <- as.Date(temp$Month)
for(i in 3:ncol(temp)){
  if(mean(temp[[i]],na.rm=T)>10000000){
    temp[[i]] <- temp[[i]]/1000000
    colnames(temp)[i] <- paste0(colnames(temp)[i],'(MM)')
  } else if(mean(temp[[i]],na.rm=T)>10000){
        temp[[i]] <- temp[[i]]/1000
    colnames(temp)[i] <- paste0(colnames(temp)[i],'(M)')
  }
}

```

---

# Data Review - Fenbid Forecasting Report

**Nielsen Advanced Analytic Consulting**

`r Sys.Date()`

---

## Content


**1 Data summary**

**2 Base Factors**

-- 2.1 Weighted Distribution

-- 2.2 DU Price

-- 2.3 Category Value

**3 ATL Factors**

-- 3.1 TV

-- 3.2 OTV 

-- 3.3 Digital

-- 3.4 SEM 

**4 BTL Factors**

-- 4.1 Brand Reminder 

-- 4.2 POSM 

<br/>

---

### 1.Data summary

<br/>

```{r echo=FALSE, warning=FALSE, message=FALSE}
options(digits=4,scipen=200)
X <- data.table(
  select(temp,1,2),
  select(temp,5,6,7),
  select(temp,8:13)
)

X.summary <- melt(select(X,-Status),id='Month') %>%
  group_by(year=year(Month),variable) %>%
  summarise(value=sum(value)) %>% arrange(variable) %>%
  mutate(value=ifelse(variable=='Weighted.Distribution',value/12,value)) %>%
  reshape2::acast(variable~year) %>% round(1)
kable(X.summary)%>%kable_styling(bootstrap_options = c("striped", "hover"))
```

<br/>

---

<br/>

### 2. Volumetric Data Review

<br/>

#### 2.1 Weighted Distribution


```{r echo=FALSE, warning=FALSE, message=FALSE}
p <- melt(select(X,1:3),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))
p1 <- ggplot() + geom_line(data=p,aes(x=Month,y=value,colour=Status,group=Status),size=1) +
  labs(x='Month',y=unique(p$variable),colour='')+ theme(legend.position="top")
p <- p %>% group_by(Month=year(Month),variable) %>% summarise(value=mean(value))
p2 <- ggplot() + geom_bar(data=p,aes(x=Month,y=value,fill=c(rep('1',3),'2')),stat='identity') +
  labs(x='Year',y=unique(p$variable)) + theme(legend.position="none")

```

`r paste0(unique(p$variable),' for Fenbid is expected at about ',round(p$value[4],2),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`

```{r echo=FALSE, warning=FALSE, message=FALSE}
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))

print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

p <- melt(select(X,1:3),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))%>%filter(year(Month)>=year(Sys.Date())-1)%>% group_by(year=year(Month),Month=month(Month),variable) %>% summarise(value=mean(value))
p$value <- round(p$value,2)

kable(reshape2::acast(p,year~Month))%>%kable_styling(bootstrap_options = c("striped", "hover"))


```

<br/>

<br/>

#### 2.2 DU Price



```{r echo=FALSE, warning=FALSE, message=FALSE}         


p <- melt(select(X,c(1:2,4)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))
p1 <- ggplot() + geom_line(data=p,aes(x=Month,y=value,colour=Status),size=1) +
  labs(x='Month',y=unique(p$variable),colour='')+ theme(legend.position="top")
p <- p %>% group_by(Month=year(Month),variable) %>% summarise(value=mean(value))
p2 <- ggplot() + geom_bar(data=p,aes(x=Month,y=value,fill=c(rep('1',3),'2')),stat='identity') +
  labs(x='Year',y=unique(p$variable)) + theme(legend.position="none")



```
`r paste0(unique(p$variable),' for Fenbid is expected at about ',round(p$value[4],2),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`
```{r echo=FALSE, warning=FALSE, message=FALSE}         
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))

print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

p <- melt(select(X,c(1:2,4)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))%>%filter(year(Month)>=year(Sys.Date())-1)%>% group_by(year=year(Month),Month=month(Month),variable) %>% summarise(value=mean(value))
p$value <- round(p$value,2)

kable(reshape2::acast(p,year~Month))%>%kable_styling(bootstrap_options = c("striped", "hover"))

```

<br/>

<br/>

#### 2.3 Category Value




```{r echo=FALSE, warning=FALSE, message=FALSE} 

p <- melt(select(X,c(1:2,5)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))
p1 <- ggplot() + geom_line(data=p,aes(x=Month,y=value,colour=Status),size=1) +
  labs(x='Month',y=unique(p$variable),colour='')+ theme(legend.position="top")
p <- p %>% group_by(Month=year(Month),variable) %>% summarise(value=sum(value))
p2 <- ggplot() + geom_bar(data=p,aes(x=Month,y=value,fill=c(rep('1',3),'2')),stat='identity') +
  labs(x='Year',y=unique(p$variable)) + theme(legend.position="none")


```
`r paste0(unique(p$variable),' for Fenbid is expected at about ',comval(p$value[4]),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`
```{r echo=FALSE, warning=FALSE, message=FALSE} 

grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))
print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))
p <- melt(select(X,c(1:2,5)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))%>%filter(year(Month)>=year(Sys.Date())-1)%>% group_by(year=year(Month),Month=month(Month),variable) %>% summarise(value=mean(value))
p$value <- round(p$value,2)
kable(reshape2::acast(p,year~Month))%>%kable_styling(bootstrap_options = c("striped", "hover"))
```

<br/>

---

<br/>

### 3 ATL Data Review

<br/>

#### 3.1 TV Spending



```{r echo=FALSE, warning=FALSE, message=FALSE}

p <- melt(select(X,c(1:2,6)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))
p1 <- ggplot() + geom_col(data=p,aes(x=Month,y=value,fill=Status),size=1,stat='identity') +
  labs(x='Month',y=unique(p$variable),fill='')+ theme(legend.position="top")
p <- p %>% group_by(Month=year(Month),variable) %>% summarise(value=sum(value))
p2 <- ggplot() + geom_bar(data=p,aes(x=Month,y=value,fill=c(rep('1',3),'2')),stat='identity') +
  labs(x='Year',y=unique(p$variable)) + theme(legend.position="none")

```
`r paste0(unique(p$variable),' for Fenbid is expected at about ',comval(p$value[4]),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`
```{r echo=FALSE, warning=FALSE, message=FALSE}
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))

print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

p <- melt(select(X,c(1:2,6)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))%>%filter(year(Month)>=year(Sys.Date())-1)%>% group_by(year=year(Month),Month=month(Month),variable) %>% summarise(value=mean(value))
p$value <- round(p$value,2)
kable(reshape2::acast(p,year~Month))%>%kable_styling(bootstrap_options = c("striped", "hover"))
```

<br/>

<br/>

#### 3.2 OTV Spending



```{r echo=FALSE, warning=FALSE, message=FALSE}         

p <- melt(select(X,c(1:2,7)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))
p1 <- ggplot() + geom_col(data=p,aes(x=Month,y=value,fill=Status),size=1,stat='identity') +
  labs(x='Month',y=unique(p$variable),fill='')+ theme(legend.position="top")
p <- p %>% group_by(Month=year(Month),variable) %>% summarise(value=sum(value))
p2 <- ggplot() + geom_bar(data=p,aes(x=Month,y=value,fill=c(rep('1',3),'2')),stat='identity') +
  labs(x='Year',y=unique(p$variable)) + theme(legend.position="none")


```
`r paste0(unique(p$variable),' for Fenbid is expected at about ',comval(p$value[4]),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`
```{r echo=FALSE, warning=FALSE, message=FALSE}         
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))

print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

p <- melt(select(X,c(1:2,7)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))%>%filter(year(Month)>=year(Sys.Date())-1)%>% group_by(year=year(Month),Month=month(Month),variable) %>% summarise(value=mean(value))
kable(reshape2::acast(p,year~Month))%>%kable_styling(bootstrap_options = c("striped", "hover"))
```

<br/>

<br/>

#### 3.3 Digital Spending



```{r echo=FALSE, warning=FALSE, message=FALSE}     

p <- melt(select(X,c(1:2,8)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))
p1 <- ggplot() + geom_col(data=p,aes(x=Month,y=value,fill=Status),size=1,stat='identity') +
  labs(x='Month',y=unique(p$variable),fill='')+ theme(legend.position="top")
p <- p %>% group_by(Month=year(Month),variable) %>% summarise(value=sum(value))
p2 <- ggplot() + geom_bar(data=p,aes(x=Month,y=value,fill=c(rep('1',3),'2')),stat='identity') +
  labs(x='Year',y=unique(p$variable)) + theme(legend.position="none")

```
`r paste0(unique(p$variable),' for Fenbid is expected at about ',comval(p$value[4]),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`
```{r echo=FALSE, warning=FALSE, message=FALSE}     
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))

print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))
p <- melt(select(X,c(1:2,8)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))%>%filter(year(Month)>=year(Sys.Date())-1)%>% group_by(year=year(Month),Month=month(Month),variable) %>% summarise(value=mean(value))
kable(reshape2::acast(p,year~Month))%>%kable_styling(bootstrap_options = c("striped", "hover"))
```


<br/>

<br/>

#### 3.4 SEM Spending



```{r echo=FALSE, warning=FALSE, message=FALSE}     

p <- melt(select(X,c(1:2,9)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))
p1 <- ggplot() + geom_col(data=p,aes(x=Month,y=value,fill=Status),size=1,stat='identity') +
  labs(x='Month',y=unique(p$variable),fill='')+ theme(legend.position="top")
p <- p %>% group_by(Month=year(Month),variable) %>% summarise(value=sum(value))
p2 <- ggplot() + geom_bar(data=p,aes(x=Month,y=value,fill=c(rep('1',3),'2')),stat='identity') +
  labs(x='Year',y=unique(p$variable)) + theme(legend.position="none")

```
`r paste0(unique(p$variable),' for Fenbid is expected at about ',comval(p$value[4]),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')``

```{r echo=FALSE, warning=FALSE, message=FALSE}     
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))
print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))
p <- melt(select(X,c(1:2,9)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))%>%filter(year(Month)>=year(Sys.Date())-1)%>% group_by(year=year(Month),Month=month(Month),variable) %>% summarise(value=mean(value))
kable(reshape2::acast(p,year~Month))%>%kable_styling(bootstrap_options = c("striped", "hover"))
```

<br/>

---

<br/>

### 4 BTL Data Review

<br/>

#### 4.1 Brand Reminder Spending



```{r echo=FALSE, warning=FALSE, message=FALSE}

p <- melt(select(X,c(1:2,10)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))
p1 <- ggplot() + geom_col(data=p,aes(x=Month,y=value,fill=Status),size=1,stat='identity') +
  labs(x='Month',y=unique(p$variable),fill='')+ theme(legend.position="top")
p <- p %>% group_by(Month=year(Month),variable) %>% summarise(value=sum(value))
p2 <- ggplot() + geom_bar(data=p,aes(x=Month,y=value,fill=c(rep('1',3),'2')),stat='identity') +
  labs(x='Year',y=unique(p$variable)) + theme(legend.position="none")

```
`r paste0(unique(p$variable),' for Fenbid is expected at about ',comval(p$value[4]),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`
```{r echo=FALSE, warning=FALSE, message=FALSE}
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))

print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

p <- melt(select(X,c(1:2,10)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))%>%filter(year(Month)>=year(Sys.Date())-1)%>% group_by(year=year(Month),Month=month(Month),variable) %>% summarise(value=mean(value))
kable(reshape2::acast(p,year~Month))%>%kable_styling(bootstrap_options = c("striped", "hover"))
```

<br/>

<br/>

#### 4.2 POSM Spending



```{r echo=FALSE, warning=FALSE, message=FALSE}

p <- melt(select(X,c(1:2,11)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))
p1 <- ggplot() + geom_col(data=p,aes(x=Month,y=value,fill=Status),size=1,stat='identity') +
  labs(x='Month',y=unique(p$variable),fill='')+ theme(legend.position="top")
p <- p %>% group_by(Month=year(Month),variable) %>% summarise(value=sum(value))
p2 <- ggplot() + geom_bar(data=p,aes(x=Month,y=value,fill=c(rep('1',3),'2')),stat='identity') +
  labs(x='Year',y=unique(p$variable)) + theme(legend.position="none")

```
`r paste0(unique(p$variable),' for Fenbid is expected at about ',comval(p$value[4]),' in CY2020,')`
`r paste0('which is ',round(abs(round(p$value[4]/p$value[3]*100,2)-100),2),'% ', ifelse(p$value[4]/p$value[3]>1,'higher ','lower '),'than the value in 2019.\n')`

```{r echo=FALSE, warning=FALSE, message=FALSE}
grid.newpage()
pushViewport(viewport(layout = grid.layout(5,6)))

print(p1, vp = vplayout(1:3,1:6))
print(p2, vp = vplayout(4:5,2:5))

p <- melt(select(X,c(1:2,10)),id=c('Month','Status')) %>% mutate(variable=gsub('\\.',' ',variable))%>%filter(year(Month)>=year(Sys.Date())-1)%>% group_by(year=year(Month),Month=month(Month),variable) %>% summarise(value=mean(value))
kable(reshape2::acast(p,year~Month))%>%kable_styling(bootstrap_options = c("striped", "hover"))
```

<br/>

---

<br/>

### END of Data Review
